### 2022-03-30 manuscript/Preface/abstract.tex
ISU thesis format CC all contained in manuscript folder

\specialchapt{ABSTRACT}

Collaboration among multiple teams has played a major role in probabilistic
forecasting of influenza outbreaks, the COVID-19 pandemic, other disease
outbreaks and problems in many other fields. When collecting forecasts from 
individual teams, ensuring that each team's model represents forecast
uncertainty according to the same format allows for direct comparison of 
forecasts as well as methods of constructing multi-model ensemble forecasts.
This paper outlines several common probabilistic forecast formats including
probability densities, samples, binned distributions and  quantile 
forecasts and compares
their usage within the context of collaborative projects. We propose the use of
a discrete mixture distribution format in collaborative forecasting in place of
other formats. The flexibility in distribution shape, the ease for scoring and 
building ensemble models and the reasonably low level of computer storage 
required to save such a format make the discrete mixture distribution an 
attractive alternative to the other mentioned formats.