### 2022-03-30 manuscript/Preface/acknowl.tex

\specialchapt{ACKNOWLEDGMENTS}

% I would like to take this opportunity to express my thanks to those
% who helped me with various aspects of conducting research and the writing
% of this thesis.
% First and foremost, Dr. Susan D. Ross for her guidance, patience and support
% throughout this research and the writing of this thesis.
% Her insights and words of encouragement have often inspired me and renewed
% my hopes for completing my graduate education.
% I would also like to thank my committee members for their efforts
% and contributions to this work: Dr. August Tanner and
% Dr. Lewis Hargrave.
% I would additionally like to thank
% Dr. Tanner for his guidance throughout the initial stages of my
% graduate career and Dr. Hargrave for his inspirational teaching style.