### 2022-04-05 manuscript/Preface/acknowl.tex

\specialchapt{ACKNOWLEDGMENTS}

I want to thank Dr. Jarad Niemi for all he did to teach and mentor me while 
writing this
creative component. His patience, scientific insights, and attention to detail
have greatly enhanced my vision and appreciation for statistics and academic 
research. I 
also want to thank Dr. Karin Dorman and Dr. Kori Khan for serving on my 
committee.

% I would like to take this opportunity to express my thanks to those
% who helped me with various aspects of conducting research and the writing
% of this thesis.
% First and foremost, Dr. Susan D. Ross for her guidance, patience and support
% throughout this research and the writing of this thesis.
% Her insights and words of encouragement have often inspired me and renewed
% my hopes for completing my graduate education.
% I would also like to thank my committee members for their efforts
% and contributions to this work: Dr. August Tanner and
% Dr. Lewis Hargrave.
% I would additionally like to thank
% Dr. Tanner for his guidance throughout the initial stages of my
% graduate career and Dr. Hargrave for his inspirational teaching style.